### 2018-02-24 paper/main.tex
jnw appendix edits

\documentclass[12pt,modern]{aastex61}
\usepackage{graphics,graphicx}
\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{comment}
\usepackage{grffile} % checks for multiple dots in figure filenames
\usepackage{empheq} % for boxing equations
\newcommand*\widefbox[1]{\fbox{\hspace{0.2cm}#1\hspace{0.2cm}}}

%% Reintroduced the \received and \accepted commands from AASTeX v5.2
%\received{July 1, 2016}
%\revised{September 27, 2016}
%\accepted{\today}
%% Command to document which AAS Journal the manuscript was submitted to.
%% Adds "Submitted to " the arguement.
\submitjournal{AAS journals.}

\shortauthors{Bouma, Masuda, Winn}
\shorttitle{Binarity and planet occurrence rates}

\newcommand{\pp}{\mathcal{P}}
\newcommand{\ps}{\mathcal{S}}
\renewcommand{\a}{_{\rm a}}
\newcommand{\s}{_{\rm s}}
\newcommand{\p}{_{\rm p}}
\renewcommand{\d}{_{\rm d}}
\renewcommand{\b}{_{\rm b}}
\begin{document}

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
  Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the size the errors are potentially larger: the occurrence of
Earth-sized planets could be overestimated by as much as 50\%. We also
show that whenever high-resolution imaging reveals a transit host star
to be a binary, the planet is usually more likely to orbit the primary
star than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}

One of the goals of exoplanetary science is to establish how common,
or rare, are planets of various types.  Knowledge of planet occurrence
rates is helpful for inspiring and testing theories of planet
formation, designing the next generation of planet-finding surveys,
and simply satisfying our curiosity.  One method for measuring
occurrence rates is to monitor the brightnesses of many stars over a
wide field, seeking evidence for planetary transits.  This was the
highest priority of the NASA {\it Kepler} mission.  Great strides have
been made in the analysis of {\it Kepler} data, including progress
towards measuring the fraction of Sun-like stars that harbor
Earth-like planets~\citep{
  youdin_exoplanet_2011,petigura_prevalence_2013,dong_fast_2013,
  foreman-mackey_exoplanet_2014,burke_terrestrial_2015}.

A lingering concern about these studies is that in most cases,
investigators have assumed that all of the targets in the survey
are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the sources that are monitored in a transit survey are
unresolved multiple-star systems, mainly binaries.  Unrecognized
binaries cause numerous systematic errors in the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, a binary system presents two opportunities to detect
transiting planets, which could increase the overall number of
detections.

At the outset of this study it was not clear to us whether the neglect
of binaries is a serious problem, or even whether the net effect of
the errors is positive or negative.  The goals of this study were to
provide a framework for dealing with these issues, and to gauge at
least the order of magnitude of the systematic errors.  In this
spirit, our models are idealized.  We do not attempt a detailed
correction of the results from {\it Kepler} or any other real transit
survey.  We took inspiration from the insightful analytic model for
transit surveys by~\citet{pepper_using_2003}.

This paper is organized as follows.  The next section enumerates the
various errors that arise from unrecognized binaries.  Then in
Section~\ref{sec:simplest}, we develop an idealized model of a transit
survey in which all planets have identical properties, and all stars
are identical except that some fraction are in binary systems.  This
simple model motivates the derivation of a general formula, given in
Section~\ref{sec:general_formula}, that allows for more realistic
stellar and planetary populations.  We use this formula in
Section~\ref{sec:more_complicated} to explore more realistic models.
We discuss the errors due to unrecognized binaries for specific cases
of current interest: the occurrence of Earth-like planets; the
apparent discrepancy between hot Jupiter occurrence rates measured in
different surveys; and the shape of the ``evaporation valley'' in the
planet radius distribution that was brought to light by
\citet{fulton_california-_2017}.  We summarize and discuss all the
results in Section~\ref{sec:discussion}.

\section{Understanding the errors}
\label{sec:concept}

Imagine that a group of astronomers wants to measure the mean number
of planets per star.  They are particularly interested in planets of
radius $r$ and stars of mass $M$ and radius $R$.  They obtain a time
series of images of some region of the sky, and prepare light curves
for a large number of unresolved sources.  Then they search these
light curves for transit signals and detect all the signals for which
\begin{equation}
  \frac{\delta}{\sigma}
  >
  \left(\frac{{\rm S}}{{\rm N}}\right)_{\rm min}.
\label{eq:S_N_thresh}
\end{equation}
Here the signal, $\delta$, is the observed transit depth, the
dimensionless fraction by which the total light fades during transits.
Note that although $\delta$ is often equated with $(r/R)^2$, this is
not true when the host star is a member of an unresolved binary.  In
those cases, $\delta$ is smaller than $(r/R)^2$ because of the
constant light from the binary companion, an effect often called the
``dilution'' of the transit signal.  The noise, $\sigma$, is the
fractional uncertainty in the determination of the flux of the source,
which may include multiple stars that are blended together.  The
threshold signal-to-noise ratio depends on the desired level of
confidence that the signal is real.
% The transit duration also affects detectability, but we omit it in
% this work for brevity.

The astronomers analyze their data assuming that all the sources are
single stars.  In particular they do not have accurate enough
parallaxes to tell that some of the stars appear to be overluminous.
They count the number $N_{\rm det}$ of transit signals that appear to
be produced by the desired type of planet around the desired type of
star.  They also count the number $N_\star$ of ``searchable stars'' in
their survey, i.e., the number of stars of the desired type that are
bright enough to have allowed for the detection of a transit signal
with amplitude $(r/R)^2$.  They estimate the occurrence rate to be
\begin{equation}
  \Lambda = \frac{N_{\rm det}}{N_\star}
                    \, \frac{1}{p_{\rm tra}},
  \label{eq:occ_rate_simple}
\end{equation}
where the geometric transit probability, $p_{\rm tra}$, accounts for
the fact that most planetary orbits are not aligned close enough with
our line of sight to produce transits.

There are many potential pitfalls in this calculation.  Some genuine
transit signals are missed even if they formally exceed the
signal-to-noise threshold, because of the probabilistic nature of
transit detection.  Planets can be misclassified due to statistical
and systematic errors in the catalogued properties of the stars.  Some
transit-like signals are spurious, arising from noise fluctuations or
failures of ``detrending'' the astrophysical or instrumental
variations in the photometric signal.  Poor angular resolution leads
to blends between eclipsing binary stars and other stars along nearly
the same line of sight, producing signals that mimic those of
transiting planets.

Here, though, we will focus exclusively on problems that arise from
the fact that many stars exist in gravitationally bound binary
systems.  We will also focus on the errors in planet occurrence as a
function of radius, rather than orbital period. This is because when
more than one transit is detected (as is usually required by the
surveyors), the orbital periods can be measured without ambiguity
regardless of whether the host star is single or one member of a
binary.  Even with this narrow focus, there are numerous sources of
error.  All three of the quantities in
Equation~\ref{eq:occ_rate_simple} are biased:
\begin{enumerate}
%    
    \item The number of detected planets, $N_{\det}$, is actually the
      number of detected planets that {\it appear} to have size $r$,
      orbiting stars that {\it appear} to have mass $M$.  Whenever the
      planet-hosting star is part of a binary,
%    
    \begin{itemize}
        \item the planet's size could be misclassified because of the
          reduction in the amplitude of the photometric signal;
%        
        \item the host star's properties could be misclassified
          because its light is combined with a second star of a
          different spectral type.
%        
    \end{itemize}
%    
    \item The number of searchable stars, $N_\star$, is biased
%    
    \begin{itemize}
%        
        \item toward lower values, because it does not include all of
          the secondary stars that were inadvertently searched for
          transiting planets;
%        
        \item toward higher values, because some of the stars that
          appeared to be searchable are in fact binaries for which the
          amplitude of the photometric signal would have been reduced
          to an undetectable level.
%        
    \end{itemize}
%    
    \item The transit probability $p_{\rm tra}$ is biased because the
      planet-hosting star could be misclassified.  At fixed orbital
      period, the transit probability is proportional to
      $\rho^{-1/3}$, where $\rho$ is the stellar mean
      density~\citep{Winn2010}.  Therefore, any errors in determining
      the host star's mean density lead to errors in the correction
      for the transit probability.
%    
\end{enumerate}

There are at least two other complications that may arise, which are
not represented in Equation~\ref{eq:occ_rate_simple}.  The first one
is an observational effect. Within the sample of apparently searchable
stars, the ratio between the number of binary and single stars will
differ from the ratio that would be found in a volume-limited sample.
This is due to a type of Malmquist bias.  The total luminosity of a
binary is larger than the luminosity of either the primary or
secondary star.  This means that for transit signals of a given
amplitude, sources that are binaries appear to be searchable at
greater distances from the Earth.  Binaries are therefore
over-represented in the collection of apparently searchable stars.

The other complication is astrophysical: the true occurrence rate of a
certain type of planet may depend on whether the host is a single
star, the primary star of a binary, or the secondary star of a binary.
The rate might also depend on the characteristics of the binary, such
as the mass ratio and orbital period.  Such differences could be
caused by the requirement for long-term dynamical stability, or
differences in the planet formation process.  When the search sample
includes both singles and binaries, the detected planets are thereby
drawn from different occurrence distributions~\citep[see][]{
  wang_occurrence_2015,kraus_impact_2016}.

Given all of the confusing and opposing sources of error, we will
proceed in stages.  We start with a model so simple that everything
can be written down on the back of a napkin, and build up to an
analytic model allowing for generality in the distribution of the
binaries and the planets they host.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Simple models}
\label{sec:simplest}

\subsection{One type of star, one type of planet}
\label{sec:model_1}

Since the effects of binarity are most pronounced when the two stellar
components are similar, we begin by considering a universe in which
all stars are identical, with mass $M$, radius $R$, and luminosity
$L$.  Single stars are uniformly distributed in space with a number
density of $n\s$ stars per cubic parsec, and binaries are uniformly
distributed with number density $n\b$.  In this scenario, stars are
never misclassified because the combined light of a binary has the
same color and spectrum as a single star.  We further assume that all
planets have the same radius, $r$, and occur around single stars and
members of binaries at the same rate, $\Lambda(r)$.

Our naive observers conduct a transit survey. To calculate the
occurrence rate of planets with radius $r$, they count the number of
detections of signals with amplitude $(r/R)^2$.  Then they identify
all the sources that appear to have been searchable for a signal of
amplitude $(r/R)^2$.  Here and throughout the rest of this paper, we
assume that the limiting source of noise is the photon-counting noise
from the source, i.e., $\sigma \propto 1/\sqrt{F}$, where
where $F$ is the total flux of the source.
Thus the observers determine the minimum $F_0$ for which detection would have been
possible, and count the number of sources with $F_{\rm tot}>F_0$. This will
include all the single stars out to a maximum distance
\begin{equation}
  d_0 = \sqrt{\frac{L}{4\pi F_0}}.
  \label{eq:dmax}
\end{equation}
Since binaries are twice as luminous, the condition $F>F_0$ will
include binaries out to the larger distance of $d_0\sqrt{2}$.
None of the stars in binaries will
appear to have a planet of radius $r$, because of the dilution of
the transit signal.  Thus, the {\it apparent} occurrence rate $\Lambda\a$ of
planets of radius $r$ is
\begin{equation}
    \Lambda\a(r) = 
        \frac{\Lambda(r) \, n\s d_0^3}
        {n\s d_0^3 + n\b (d_0\sqrt{2})^3}.
    \label{eq:rate_at_rp}
\end{equation}
This apparent rate is larger than the true rate by a factor
\begin{equation}
    \frac{\Lambda\a(r)}{\Lambda(r)} = 
        \frac{1}{1 + 2^{3/2}(n\b/n\s)}.
    \label{eq:correction_rp}
\end{equation}

The observers will also detect some transiting planets around stars
with binary companions.  The amplitude of these signals is
\begin{equation}
  \frac{L(r/R)^2}{L + L} = \frac{1}{2}\left(\frac{r}{R}\right)^2 =
  \left(\frac{r/\sqrt{2}}{R}\right)^2,
    \label{eq:delta_obs_general} 
\end{equation}
leading the astronomers to believe they have discovered a population
of planets with radius $r/\sqrt{2}$.  To calculate the corresponding
occurrence rate, they count the stars for which this type of signal
would have been detectable.  The limiting flux for detection in this
case is $4F_0$, because the signal amplitude hase been reduced by a
factor of two and the noise level must also be reduced by a factor of
two.  The condition $F>4F_0$ is met for single stars within a distance
$d_0/2$, and binaries within a distance $d_0\sqrt{2}/2$.  Therefore,
the observers will calculate the occurrence rate of this new type of
planet to be
\begin{equation}
    \Lambda\a\left( \frac{r}{\sqrt{2}} \right) = 
        \frac{2\Lambda(r) \, n\b (d_0\sqrt{2}/2)^3}
        {n\s (d_0/2)^3 + n\b (d_0\sqrt{2}/2)^3}
    =
    \frac{2\Lambda(r)\cdot2^{3/2}(n\b/n\s)}{1 + 2^{3/2} (n\b/n\s)}.
    \label{eq:correction_diluted_rp}
\end{equation}
Figure~\ref{fig:model_1_volumes} illustrates the
volumes enclosing the apparently searchable single and binary stars.

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=0.6\textwidth]{figures/visualize_volumes.pdf}
    \end{center}
    \caption{ {\it Top.}---Volumes within which sources appear to be
      searchable for planets of radius $r$.  Single stars are
      searchable out to a distance $d_0$, at which point they become
      too faint to allow the detection of transits.  Binaries
      are brighter and therefore appear to be searchable out to a
      larger distance.  {\it Bottom.}---Volumes within
      which stars appear to be searchable for planets of radius
      $r/\sqrt{2}$.  }
    \label{fig:model_1_volumes}
\end{figure}

\begin{figure}[!tb]
    \begin{center}
        \includegraphics[width=.6\textwidth]{figures/occ_rate_vs_radius_model_1_brokenx.pdf}
    \end{center}
    \vspace{-0.5cm}
    \caption{ Apparent occurrence rate $\Lambda\a$, and true
    occurrence rate $\Lambda$ in a universe with only one type of
    star, one type of planet (with radius $R_\oplus$), and a number
    density of binary stars 5\% the number density of single stars.
    The occurrence
    rate of planets with radius $R_\oplus$ is underestimated, and the
    occurrence rate of planets with radius
    $R_\oplus/\sqrt{2}$ is overestimated.}
    \label{fig:occ_rate_model_1}
\end{figure}

We can now assess the severity of the errors, for a given value of the
binary-to-single ratio $n\b/n\s$. For stars with masses from 0.7 to
1.3~$M_\odot$, \citet{raghavan_survey_2010} found the multiplicity
fraction~--~the fraction of systems in a volume-limited sample that
are multiple~--~to be 0.44.  Assuming all multiple systems are
binaries, this gives a binary fraction
\begin{equation}
  \frac{n\b}{n\s + n\b} \approx 0.44,
\end{equation}
which implies $n\b/n\s \approx 0.79$.  Of course not all of
these binaries are ``twin'' binaries as we have assumed in our simple
calculation.  Later, in Section~\ref{sec:general_formula}, we will
allow for a continuum of properties for the secondary stars. For now,
we might guess that only a tenth of the binaries have pairs of stars
close enough in brightness to produce errors as significant as those
we have been considering.  Adopting the value $n\b/n\s\approx0.05$, we
find
\begin{equation}
    \frac{\Lambda\a(r)}{\Lambda(r)} = 0.88,~~
    \frac{\Lambda\a(r/\sqrt{2})}{\Lambda(r)} = 0.25.
    \label{eq:ratios}
\end{equation}
All together, the various effects produce biases of order 10\% in the
occurrence rates.  We will see that this level of error is
characteristic of many of our more complicated models as well.

\subsection{Planets of different sizes}
\label{sec:model_1_density}

We will now generalize to allow for a continuum of planet sizes.   We
introduce the occurrence rate density
\begin{equation}
    \Gamma(r) \equiv \frac{{\rm d}\Lambda}{{\rm d}r},
    \label{eq:rate_density_defn}
\end{equation}
the number of planets per star per unit planet radius.  The naive
observers are measuring $\Gamma\a(r\a)$, the apparent occurrence rate
density of planets with an ``apparent radius''
\begin{equation}
  r\a = R \sqrt{\delta}.
\end{equation}

For a given value of $r\a$, the observers assemble the sample of
$N_\star(r\a)$ sources for which a signal of amplitude $\delta =
(r\a/R)^2$ could have been detected.  We showed in the previous
section that when all stars are identical, this sample will contain
binaries and single stars in the ratio $2^{3/2}(n\b/n\s)$.  Here we
will denote this ratio by $\mu$, which will be calculated later in a
more general context.  With this definition, the number of single
stars is proportional to $1/(1+\mu)$ and the number of binaries is
proportional to $\mu/(1+\mu)$.

From within this sample of apparently searchable sources, the
observers sum the number of detections $dN_{\rm det}$ of planets
with an apparent radius between $r\a$ and $r\a + dr\a$.  There will be
three contributions to this sum,
\begin{align}
  {\rm d}N_{{\rm det},0} &= 
      \frac{N_\star(r\a) p_{\rm tra}}{1+\mu}
      \Gamma_0(r\a)\mathrm{d}r\a,
  \label{eq:n0} \\
  {\rm d}N_{{\rm det},1} &= 
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_1(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a),
	\label{eq:n1} \\
  {\rm d}N_{{\rm det},2} &=
      \frac{\mu N_\star(r\a) p_{\rm tra}}{1+\mu} 
      \Gamma_2(\sqrt{2}r\a)\mathrm{d}(\sqrt{2}r\a).
	\label{eq:n2}
\end{align}
where here and elsewhere, the subscript 0 refers to single stars, 1
refers to primary stars of binaries, and 2 refers to secondary stars
of binaries. The apparent occurrence rate density is therefore
\begin{equation}
  \Gamma\a(r\a) = \frac{{\rm d}N_{\rm det}}{{\rm d}r\a} \frac{1}{N_\star(r\a) p_{\rm tra}} =
  \frac{\Gamma_0(r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_1(\sqrt{2}r\a)}{1+\mu} +
  \frac{\mu\sqrt{2}\,\Gamma_2(\sqrt{2}r\a)}{1+\mu}.
	\label{eq:formula_twin_binary}
\end{equation}
Our earlier results, Equations~\ref{eq:rate_at_rp}
and~\ref{eq:correction_diluted_rp}, are recovered by integrating this formula
over $r$ after inserting
\begin{equation}
  \Gamma_i(r) = \Lambda(r\p)\, \hat{\delta}(r-r\p),~~{\rm for}\ i\in\{0,1,2\}
\end{equation}
where $\hat{\delta}$ is the Dirac delta function, and $r\p$ is the
single planet size considered in Section~\ref{sec:model_1}.

\subsection{Binaries with different mass ratios}
\label{sec:general_formula}

Next we generalize to allow for a spectrum of different properties for
the secondary stars.  For simplicity we assume that the stars form a
one-parameter family specified by the stellar mass $M$.  This is
approximately the case for main-sequence stars.  The functions $L(M)$
and $R(M)$ give the luminosity and radius as a function of mass, and
$f(q)$ is the distribution of binary mass ratios in a volume-limited
sample.  We also assume that observers perceive all the binaries to be
isolated stars with the same mass as the primary star, i.e., the light
from the secondary star is either too faint or too similar to the
primary star to make a difference in the stellar classification.
Finally, for clarity of presentation we make the simplifying
assumption that $\Gamma_0$, $\Gamma_1$ and $\Gamma_2$ do not depend on
stellar mass, although our formalism can easily accommodate such a
dependence.

To compute the apparent occurrence rate density, we need to make
the following modifications to Equation~\ref{eq:formula_twin_binary}:
%
\begin{enumerate}
%
\item The Malmquist bias is different.  For a given value of $r\a$, the
  number of binary systems in the searchable sample with mass ratio
  $(q, q+\mathrm{d}q)$ is
\begin{equation}
    {N_\star(r\a)\over{1+\mu}}\, {n_{\rm b}\over n_{\rm s}}
    \left[{L(M) + L(qM) \over L(M)}\right]^{3/2}
    f(q)\,\mathrm{d}q,
    \label{eq:N_binaries}
\end{equation}
where $\mu$ is given by
\begin{equation}
    \mu= 
    %\frac{N_\star^{\rm b}(r\a)}{N_s^0(r\a)}=
    \int_0^1 {n_{\rm b}\over n_{\rm s}}\left[{L(M) + L(qM)
    \over L(M\a)}\right]^{3/2} f(q)\,\mathrm{d}q.
    \label{eq:mu_general}
\end{equation}
To obtain the total number of searchable binaries,
Equation~\ref{eq:N_binaries} needs to be integrated over $q$. The
result of the integration replaces ${\mu N_\star(r\a)/(1+\mu)}$ in
Equations~\ref{eq:n1} and~\ref{eq:n2}.

\item The apparent radius of a planet in a binary system now depends
on whether the host star is the primary or the secondary star. When the host
is the primary star, we write $r=\mathcal{D}_1 r\a$, where
\begin{equation}
    \mathcal{D}_1
    = \left[ \frac{ L(M) + L(qM) }{ L(M) } \right]^{1/2}
    \label{eq:dil_1}
\end{equation}
is the appropriate dilution factor.
When the host is the secondary star, a correction must also be made
to account for the different radius of the secondary star.
In that case $r=\mathcal{D}_2 r\a$, where
\begin{equation}
  \mathcal{D}_2
  = \frac{R(qM)}{R(M)}
    \left[ \frac{L(M) + L(qM)}{L(qM)} \right]^{1/2}.
  \label{eq:dil_2}
\end{equation}

\item When a transiting planet is detected around a secondary star,
  the naive observers make the wrong correction for the transit
  probability.  At a fixed orbital period, $p_{\rm tra}$
  in Equation~\ref{eq:n2} must be multiplied by a factor of
\begin{equation}
  {R(qM) \over R(M)}q^{-1/3}.
\end{equation}

\end{enumerate}

Taking these modifications into account, a general formula for the
apparent rate density is
\begin{align}
    \notag
    \Gamma\a(r\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a)\cdot
         \frac{R(qM)}{R(M)} q^{-1/3}
    \right] \right\}.
    \label{eq:general_Gamma_a}
\end{align}
The \hyperref[sec:appendix]{appendix} gives an alternative derivation
of this equation.

We note that the combination $\mathcal{D}_1^3f(q)$ is the mass-ratio
distribution for binaries contained within the searchable sample of
sources.  With this in mind we may rewrite
Equation~\ref{eq:general_Gamma_a} as
\begin{align}
    %\notag
    %\notag
    \Gamma\a(r\a)
    =%&= 
    {1\over 1+\mu} \,
    \left[
       \Gamma_0(r\a)%\phantom{\tilde \Gamma}\right.%&\\
       %&\left.
       +\mu\,
       \left\langle
       \mathcal{D}_1\cdot\Gamma_1\left(\mathcal{D}_1r\a\right)
       +
       {q \mathcal{D}_2}\cdot\Gamma_2\left(\mathcal{D}_2r\a\right)
       \cdot
       {R(qM) \over R(M)}q^{-1/3}
       \right\rangle
    \right],
\end{align}
where the angle brackets denote averaging over all the binaries in the
searchable sample.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Case Studies}
\label{sec:more_complicated}

We can now gauge the size of the systematic errors associated
with unresolved binaries, for cases of interest.  Throughout this
section we assume $R\propto M$, and $L \propto M^{\alpha}$ where
$\alpha = 3.5$, as is roughly the case for main-sequence stars.
With these assumptions, Equations~\ref{eq:dil_1} and~\ref{eq:dil_2}
become
\begin{equation}
  \mathcal{D}_1
  =  (1+q^\alpha)^{1/2}~~{\rm and}~~\mathcal{D}_2
  = q(1+q^{-\alpha})^{1/2}.
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Power-law planet radius distribution}
\label{sec:model_2}

Based on {\it Kepler} data, \citet{howard_planet_2012} found the
radius distribution of planets between 2 and 17~$R_\oplus$ orbiting Sun-like
stars within 0.25\, AU to be consistent with a power law,
\begin{equation}
  \Gamma_0(r) \propto r^\gamma,
\label{eqn:power_law}
\end{equation}
with $\gamma=-2.92\pm 0.11$.  Their analysis ignored the effects of
binarity.  We can use our formalism to estimate the resulting level of
systematic error.

To warm up we will again consider the case in which all stars are
identical.  We assume further that the distributions of planets around
primaries and secondaries differ only by multiplicative factors,
\begin{equation}
    \Gamma_1 = Z_1 \Gamma_0,~~\Gamma_2 = Z_2 \Gamma_0.
\end{equation}
Application of Equation~\ref{eq:general_Gamma_a} gives
\begin{equation}
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} = 
    \frac{1}{1+\mu}
    +
    2^{\frac{\gamma+1}{2} } \frac{\mu}{1+\mu} \left(Z_1 + Z_2\right),
    \label{eq:model5_apparent_rate_density}
\end{equation}
where $\mu = 2^{3/2} n\b/n\s$.  The ratio does not depend on $r\a$ and
thus, under our assumptions, the effect of twin binaries is simply to
change the normalization of the radius distribution.  If we further
assume that planet occurrence is independent of system multiplicity,
i.e., $Z_1=Z_2=1$, then
\begin{equation}
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} = 
    \frac{1 + 2^{\frac{\gamma+3}{2}}\mu}{1 + \mu}.
    \label{eq:power_law_correction}
\end{equation}
Adopting $\gamma=-2.92$, and a twin binary fraction $n\b/n\s=0.05$, as
before, we obtain $\mu = 0.14$ and $\Gamma\a/\Gamma_0 = 1.003$.  The
correction is tiny because, by coincidence, the reported value of
$\gamma$ is very close to $-3$, the value for which
Equation~\ref{eq:power_law_correction} reduces to unity and the
effects of binarity cancel out completely.  This suggests that
although \citet{howard_planet_2012} derived the occurrence rate
density under the false assumption that all stars are single, the
resulting systematic error is negligible.

\begin{figure}[!tb]
    \centering
    \includegraphics[width=0.6\textwidth]{figures/mass_ratio_distribution.pdf}
    \caption{ The mass ratio distribution of binaries in a
      magnitude-limited sample, assuming the underlying volume-limited
      distribution is uniform and the luminosity-mass relation is
      $L\propto M^{3.5}$. For transit signals of a given amplitude,
      the sample of searchable sources is magnitude-limited, causing
      high mass ratios to be over-represented. }
    \label{fig:q_distribn_mag_limited}
\end{figure}

For a more accurate analysis we now consider a distribution of binary
mass ratios.  Studies of binaries in the local neighborhood suggest
that the distribution of mass ratios is nearly uniform between zero
and unity~\citep{raghavan_survey_2010}.  We consider the more general
possibility of a power-law dependence, $f(q) = \mathcal{N}_q q^\beta$,
where $\mathcal{N}_q$ is a normalization constant.  In this case
Equation~\ref{eq:mu_general} gives
\begin{equation}
    \mu = \frac{1}{\mathcal{N}_q} \frac{n\b}{n\s}
    \int_0^1 (1+q^\alpha)^{3/2}q^\beta\,{\rm d}q.
    \label{eq:mu_power_law_q}
\end{equation}
Figure~\ref{fig:q_distribn_mag_limited} illustrates the Malmquist
bias for the particular case of a flat distribution.

Again assuming $\Gamma_0=\Gamma_1=\Gamma_2$, the apparent occurrence
rate density is 
\begin{align}
    %see annoying_q_factor_dec22.pdf for derivation
    \notag
    %\frac{\left.\Gamma\a(r\a)\right|_{r\a= r}}{\Gamma_0(r)} 
    \frac{\Gamma\a(r\a)}{\Gamma_0(r\a)} 
    &=
    \frac{1}{1+\mu}
    \left[
      1 + \frac{1}{\mathcal{N}_q} \frac{n\b}{n\s}
    \left(
      \int_0^1 {\rm d}q\,q^\beta (1+q^\alpha)^{\frac{\gamma+4}{2}} +
      \right.
      \right. \\
      %isn't latex great?
      &\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad\quad
      \left.\left.
      \int_0^1 {\rm d}q\,q^{\beta+\gamma+\frac{5}{3}} 
      (1+q^\alpha)^{\frac{3}{2}}
      (1+q^{-\alpha})^{\frac{\gamma+1}{2}}
    \right)
    \right],
    \label{eq:powerlaw_vary_binary}
\end{align}
Adopting the realistic numerical values $n\b/n\s=0.79$, $\alpha =
3.5$, $\beta=0$, and $\gamma=-2.92$, the summed integrals in
Equation~\ref{eq:powerlaw_vary_binary} give $(\ldots)\approx 1.503$,
and $\Gamma\a/\Gamma_0 = 1.048$.  The apparent occurrence rate density
is 4.8\% higher than the true value.  We conclude that for planets around
Sun-like stars with orbits within 0.25\,AU and radii in the range
2--17\,$R_\oplus$, the systematic errors associated with binarity are
on the order of a few percent.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Broken power-law planet radius distribution}
\label{sec:model_3}

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_3_rpu_22.5_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_3_rpu_22.5_manyZs.pdf}
    \caption{ {\it Left:} Apparent occurrence rate density
      ($\Gamma\a$), compared to the true occurrence rate density
      ($\Gamma_0$) for single stars and primary stars in binaries.
      Three different cases are plotted, with different choices for
      the true occurrence rate density for secondary stars relative to
      that for single stars ($Z_2/Z_0$).
      {\it Right:} Same, but for apparent occurrence
      rate ($\Lambda\a$) within $0.5\,R_\oplus$ bins.
      In both cases the true planet
      radius distribution is specified by
      Equation~\ref{eq:model3_radius_distribution}.
    }
    \label{fig:occ_rate_model_3_log}
\end{figure}

For planets smaller than 2\,$R_\oplus$, the true occurrence rate
density is more uncertain because such planets are more difficult to
detect.  To investigate the systematic errors associated with binarity
on the inferred occurrence rate of Earth-sized planets, we need to
make plausible assumptions about the true occurrence rate density.
Obviously if the rate density continues to vary as $r^{-3}$ down to
much smaller planet sizes, the results of the preceding section will
hold.  But there is no particular reason to think this will be the
case and indeed, some investigators have concluded that the rate
density begins to level off to a constant value as $r$ decreases below
$R_\oplus$~\citep{petigura_plateau_2013}.  To investigate the
implications we consider a broken power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2R_\oplus \\
        {\rm constant} & \text{for } r\leq2R_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
In this case the integrals that appear in
Equation~\ref{eq:general_Gamma_a} are tedious to work out
analytically, leading us to evaluate them numerically.\footnote{We
refer the interested reader to our online code for performing this
computation: \url{github.com/lgbouma/binary_biases}, commit
\texttt{6cb920c}.} Figure~\ref{fig:occ_rate_model_3_log} shows the
results, again for the case $n\b/n\s=0.79$, $\alpha = 3.5$, $\beta=0$,
and $\gamma=-2.92$.

The most obvious aspect of the results is that the occurrence rate
density of planets smaller than the breakpoint radius of 2\,$R_\oplus$
is now substantially overestimated.  If secondary stars have the same
planet population as the single stars, then the occurrence rate
density of small planets is overestimated by $50\%$.  The magnitude of
the systematic error decreases if there are fewer planets around
secondaries; if secondaries host no planets at all, the apparent rate
is overestimated by only $10\%$.

The effects on larger planets are not as substantial.  To be
quantitative we consider the occurrence rate of giant planets with
$r>8\,R_\oplus$, by integrating the occurrence rate density displayed
in Figure~\ref{fig:occ_rate_model_3_log}:
\begin{equation}
    \Lambda_{\rm giant,a} =
      \int_{8\,R_\oplus}^{\infty} \Gamma\a(r\a)\,{\rm d}r\a,~~{\rm
      and}~~
    \Lambda_{\rm giant,0} =
      \int_{8\,R_\oplus}^{\infty} \Gamma_0(r)\,{\rm d}r.
\end{equation}
In this case the occurrence rate density is underestimated.  The
difference between the apparent and true rates is largest when the
secondary stars do not host any planets, and has a magnitude
$\Lambda_{{\rm g},0}/\Lambda_{\rm g,a} = 1.13$.
%occ_rate_vs_radius_model_3_withtext_Zsub2_0.00_rpu_22.5.pdf
If instead the secondary stars have half as many giant planets as single
stars, then the factor is reduced from 1.13 to 1.06.

\subsection{Which star has the planet?}
\label{subsec:whichstar}
\begin{figure}[!t]
    \centering
    \includegraphics[width=\textwidth]{figures/ndet_vs_radius_logx_model_3_fraclines_rpu_22.5_manyZs.pdf}
    \caption{ Fraction of detected planets that orbit single stars
      (solid lines), primary stars of binaries (dashed lines), and
      secondary stars of binaries (dotted lines), as a function of the
      apparent radius of the planet.  We assume that single stars and
      primaries have the same number of planets per star.  Three
      different cases are plotted, differing in the relative
      occurrence of planets around secondary stars ($Z_2/Z_0$).  The
      true radius distribution is given by
      Equation~\ref{eq:model3_radius_distribution}.}
    \label{fig:frac_model_3}
\end{figure}

Our formalism also provides a way to calculate the relative
probability that a detected planet orbits a single star, the primary
star, or a secondary star.  These can be obtained from the
generalizations of Equations~\ref{eq:n0}--\ref{eq:n2}, which are
derived in the \hyperref[sec:appendix]{appendix}.  This information
can be used to help interpret the results of a transit survey.  It is
also relevant to the fairly common situation in which transits are
detected from a source, and subsequent high-resolution imaging reveals
the source to be a multiple-star system.  The question then arises:
which star is the host of the detected planet?  Without further
observations the answer is often unclear.  To illustrate we return to
the scenario described in the previous section: a broken power-law
radius distribution with $\gamma=-2.92$, along with the choices
$n\b/n\s=0.79$, $\alpha = 3.5$, and $\beta=0$.  We also try different
assumptions for the rate of planets around secondaries, i.e.,
different choices for $Z_2/Z_0$.  We use
Equations~\ref{eq:n_det_0}--\ref{eq:n_det_2} to compute the number of
detections from single stars, primaries, and secondaries, and divide
by the total number of detections at each apparent radius.

Figure~\ref{fig:frac_model_3} shows the results.  When the apparent
radius exceeds $2\,R_\oplus$, and planets exist about secondaries at
the same rate as the primary star ($Z_2/Z_0=1$), then the planet is
$3$ times more likely to orbit the primary star than the secondary
star.  If secondaries host half as many planets as primaries
($Z_2/Z_0=0.5$), then a detected planet is $5$ times more likely to
orbit the primary star.

The situation for smaller apparent radii is more nuanced.  For
$Z_2/Z_0=0$ or 0.5, any transit signals from binaries are still always
more likely to arise from the primary star.  However, if planets exist
at the same rate in primaries and secondaries ($Z_2/Z_0=1$), then
below apparent radii of $0.4\,R_\oplus$, more of the detected planets
in binaries come from secondaries.  They are actually much larger
planets for which $\delta$ has been substantially reduced by the light
from the primary star.  As we consider apparent radii ranging from 2
to 1.4\,$R_\oplus$, the fraction of detected planets with binary
companions increases by anywhere from $6\%$~to~$12\%$, depending on
the relative occurrence of planets about primaries and secondaries.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A gap in the radius distribution}
\label{sec:further_models}

\begin{figure}[!t]
    \centering
    \epsscale{1.15}
    \plottwo{figures/int_rate_density_vs_radius_model_4_xcut_manyZs.pdf}{figures/int_occ_rate_vs_radius_model_4_xcut_manyZs.pdf}
    \caption{ {\it Left:} True and apparent occurrence rate densities,
      and {\it right:} occurrence rates, as a function of planet
      radius, using a model for the radius distribution that exhibits
      a gap between 1.5 and 2\,$R_\oplus$
      (Equation~\ref{eq:model4_radius_distribution}). }
    \label{fig:model_4}
\end{figure}

\citet{fulton_california-_2017} recently reported a ``gap'' in the
radius distribution of close-in planets around Sun-like stars, between
planet sizes of 1.5 and 2\,$R_\oplus$.  This can also be visualized as
a ``valley'' in the occurrence rate density as a function of radius
and period.  The existence of the gap has been independently
corroborated from a sample of KOIs with asteroseismically-determined
stellar parameters~\citep{van_eylen_asteroseismic_2017}.  Such a
feature had been predicted as a consequence of the gradual
photo-evaporation of the hydrogen-helium atmospheres of small rocky
planets, during the first 100~Myr of the age of the system when the
host star produces a higher flux of ultraviolet and X-ray
radiation~\citep{owen_evaporation_2017}.

The occurrence calculations that led to this discovery, like almost
all occurrence calculations, did not carefully account for
unresolved binaries.  Intuitively we expect the effect of unresolved
binaries to have a blurring effect on the radius distribution, filling
in any gaps in the radius distribution and making them appear less
empty than in reality.  To investigate the quantitative effects we
make identical assumptions as in Section~\ref{sec:model_3}, except
that the radius distribution is assumed to have a complete absence of
planets with sizes between 1.5 and 2\,$R_\oplus$:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2\,R_\oplus, \\
        0 & \text{for } 1.5\,R_\oplus < r < 2\,R_\oplus, \\
        {\rm constant} & \text{for } r\leq 1.5\,R_\oplus.
    \end{cases}
    \right.
    \label{eq:model4_radius_distribution}
\end{align}
Figure~\ref{fig:model_4} shows the resulting true and apparent occurrence
rate densities.
For the case $Z_2/Z_0=1$, unresolved binaries reduces the contrast
of the gap by nearly a factor of two, while also producing
spurious features for apparent sizes beneath 1.5\,$R_\oplus$.
Thus the gap identified by \citet{fulton_california-_2017} may be
even more devoid of planets than it appears.

%Of course, effects unrelated to
%binarity would also ``blur'' the gap in the planet radius dimension.
%In particular, the valley's period-dependence is almost certainly not
%flat \citep{van_eylen_asteroseismic_2017,owen_evaporation_2017}.  This
%means that any study measuring the gap's width or depth while
%accounting for binarity must either perform tests at fixed orbital
%period, or else marginalize over period and account for the associated
%blurring in the planet radius dimension.

\subsection{Gaussian radius distribution}

The planet population may include some special members with a distinct
radius distribution.  For example in the recent study
by~\citet{petigura_CKS_2017}, hot Jupiters appear as an island in
period-radius space, rather than as a component of a power-law
distribution extending to smaller planets and longer periods.
For this reason we test the effects of unresolved binaries
on a Gaussian radius distribution,
\begin{equation}
    f(r) \propto \exp \left[ -\frac{(r-\bar{r})^2}{2\sigma_r^2} \right],
\end{equation}
with $\bar{r} = 14\,R_\oplus$ and $\sigma_r = 2\,R_\oplus$.  As before,
we allow for different normalizations of the planet populations around
single, primary, and secondary stars: $\Gamma_i(r) = Z_i f(r)$.
Figure~\ref{fig:gaussian_HJ} shows the results for the apparent radius
distribution, $\Gamma\a(r\a)$.  As one would expect, the effect is to
smear the radius distribution toward smaller values, because some of
the hot Jupiters now appear to be smaller planets.  Less obviously,
the integrated rate of hot Jupiters is also affected.  We calculate
the apparent hot Jupiter rate by integrating the apparent rate density
for $r\a > 8\,R_\oplus$.  If secondaries host any planets at all, the
apparent rate is {\it greater} than the true rate.  For instance, when
hot Jupiters are just as common around secondary stars as single or primary
stars, then $\Lambda_{\rm HJ,a}/\Lambda_{{\rm HJ},0} = 1.23$.
%int_rate_density_vs_radius_model_7_withtext*manyZs.pdf
The apparent rate is higher because a larger number of stars were
searched than are accounted for in the rate calculation.

\begin{figure}[!tb]
    \centering
    \includegraphics[width=.6\textwidth]{figures/int_rate_density_vs_radius_model_7_rpu_22.5_manyZs.pdf}
    \caption{ True and apparent occurrence rate densities, for a
      population of planets with true radii $r$ drawn from a Gaussian distribution
      with mean $14\,R_\oplus$ and standard deviation $2\,R_\oplus$.  This
      is similar to the hot Jupiter distribution presented
      by~\citet{petigura_CKS_2017}.  }
    \label{fig:gaussian_HJ}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Summary and Discussion}
\label{sec:discussion}

Ignoring binarity introduces systematic errors to star and planet
counts in transit surveys, which in turn lead to biases in the derived
planet occurrence rates.  Thus far, occurrence rate calculations using
transit survey data have mostly ignored stellar
multiplicity~\citep[\textit{e.g.},][]{howard_planet_2012,fressin_false_2013,foreman-mackey_exoplanet_2014,dressing_occurrence_2015,burke_terrestrial_2015}.
We do not claim to have solved this problem. Our aim was to clarify
the various effects and assess the size of the errors in various
situations. We hope that our formalism will also be useful to those
investigators who do attempt more realistic calculations of occurrence
rates.

The calculations presented in Section~\ref{sec:model_2} suggest that
for planets around Sun-like stars with periods shorter than a few
months and apparent radii exceeding $2\,R_\oplus$, binarity can
basically be ignored.  The errors are only of order a few percent.
For apparent radii below $2\,R_\oplus$, nature may not be as
forgiving.  The size of the effect depends on the true radius
distribution and in particular whether it continues to vary as
$\sim$$r^{-3}$ down to 1\,$R_\oplus$ and below.  The calculations
presented in Section~\ref{sec:model_3} suggest that the apparent rates
could be overestimated by as much as 50\%.

\paragraph{Earth-like planets} For the specific problem of calculating the occurrence rate
$\eta_\oplus$ of Earth-like planets with periods of order one year,
even a 50\% systematic error would not be the dominant source of
uncertainty at the moment.  Estimates by
\citet{youdin_exoplanet_2011},~\citet{petigura_prevalence_2013},~\citet{dong_fast_2013},~\citet{foreman-mackey_exoplanet_2014},
and~\citet{burke_terrestrial_2015} have found values of $\eta_\oplus$
ranging from $0.03$ to unity, with the wide variance arising from
small-number statistics as well as a host of other systematic
errors~\citep[see][Figure~17]{burke_terrestrial_2015}.  Rapid progress
is expected soon, though, and binarity will merit closer attention in
the future.  It may also be worth investigating the possible effects
of any dependence of occurrence rate densities on the period of the
binary.  Binaries with separations $\lesssim 10\,{\rm AU}$ could
provoke dynamical instabilities, leading to fewer Earth-like planets
per
star~\citep[\textit{e.g.},][]{holman_long-term_1999,wang_influence_2014,
  kraus_impact_2016}.  This would affect transit survey measurements
of $\eta_\oplus$ beyond our rough estimate.

\paragraph{Hot Jupiters} Another frequently discussed problem
involving occurrence rates is the discrepancy between the occurrence
rates of hot Jupiters that have been measured in different surveys, as
summarized in Table~\ref{tab:hj_rates}.  In particular, the California
Planet Search (CPS), a Doppler survey, found a higher rate ($12\pm 2$
per thousand stars) than the {\it Kepler} transit survey
($5.7^{+1.4}_{-1.2}$).  It has been suggested that unresolved binaries
in the {\it Kepler} sample might be an important factor, but the
calculations presented in Section~\ref{sec:further_models} suggest
that this is not the case.  We found that the effects of unresolved
binaries are likely to result in an overestimation of the hot-Jupiter
occurrence rate, which would worsen the discrepancy, rather than
relieve it. Some other possibilities are errors in stellar
classification of the {\it Kepler} stars, as well as differences in
the metallicity of the stars in each survey.  The occurrence rate of
hot Jupiters is known to vary strongly with stellar metallicity, with
$\Lambda \propto 10^{3.4{\rm [M/H]}}$ according to a recent
study~\citep{petigura_CKS_2017}.  \citet{guo_metallicity_2017} found
the mean metallicity of the {\it Kepler} stars to be [M/H]~$=-0.045\pm
0.009$, which is only slightly lower than that of the stars in the
Doppler survey, $-0.005\pm 0.006$.  Nevertheless, given the strong
dependence of occurrence on metallicity, even a difference of
$0.04$~dex could account for a factor of $10^{3.4\cdot 0.04} = 1.37$
in the occurrence rate.  After correcting for metallicity, the
discrepancy between the {\it Kepler} and CPS occurrence rates is
reduced to the 1.7$\sigma$ level.

\paragraph{Radius corrections} Another application of our formalism is
to help in those cases in which it is not clear whether the primary or
secondary star is the source of a transit signal.  By this point
almost all of the {\it Kepler}\ Objects of Interest have been examined
with high resolution imaging~\citep{
  howell_speckle_2011,adams_adaptive_2012,adams_adaptive_2013,horch_observations_2012,
  horch_most_2014,lillo-box_multiplicity_2012,lillo-box_high-resolution_2014,dressing_adaptive_2014,
  law_robotic_2014,cartier_revision_2015,everett_high-resolution_2015,gilliland_hubble_2015,
  wang_influence_2015,wang_influence_2015-1,baranec_robo-ao_2016,ziegler_robo-ao_2017}.
The results of these programs have been summarized
by~\citet{furlan_kepler_2017}. Based on this information,
\citet{hirsch_assessing_2017} re-assessed the radius measurements of
the {\it Kepler} planets in binaries. They found that planet radii are
generally underestimated by a factor $r/r\a = 1.17$ if all planets
orbit primaries, and $r/r\a = 1.65$ if detected planets are equally
likely to orbit primaries and singles.  Our formalism provides a way
to assess the relative probabilities that a detected planet orbits the
primary or secondary star.  The calculations presented in
\S\,\ref{subsec:whichstar} suggest that in almost all cases the planet
most likely orbits the primary star.

\begin{comment}
% JNW: These 2 sections do not seem very interesting.
  {\paragraph{Fraction of detected planets with binary companions}
A related effect is that some detected planets are more likely to be
in binary systems than others.  Notably, Figure~\ref{fig:frac_model_3}
predicts that the fraction of detected planets with binary companions
should increase by $6$~to~$12\%$ going from $2R_\oplus$ to
$1.4R_\oplus$.

In \citet{ziegler_robo-ao_2017}'s recent summary of the Robo-AO KOI
survey, they reported the fraction of planet-hosting stars with
Robo-AO detected companion stars, binned over apparent planet radii.
The fraction of KOIs with nearby stars and their $1\sigma$
uncertainties are:
\begin{itemize}
    \item Earths ($r\a < 1.6R_\oplus$): $16.3 \pm 1.0\%$, from 1480
    systems.
    \item Neptunes ($1.6R_\oplus < r\a < 3.9 R_\oplus$): $13.0 \pm
    0.8\%$, from 2058 systems.
    \item Saturns ($3.9R_\oplus < r\a < 9 R_\oplus$): $13.6 \pm 2.0\%$, from 338 systems. 
    \item Jupiters ($r\a > 9R_\oplus$): $19.0 \pm 2.8\%$, from 247 systems. 
\end{itemize} 
The uncertainties were calculated
following~\citet{burgasser_binarity_2003}.  The absolute values of the
companion fractions are lower than in the solar neighborhood at
least in part because of Robo-AO's sensitivity
(\citealt{ziegler_robo-ao_2017}'s
Figure~2;~\citealt{raghavan_survey_2010}).
%Another reason for the lower measured companion fraction could be
%that planetary systems are more rare in binary systems,
%as~\citealt{kraus_impact_2016} found for close binary companions).
The uptick in the companion rate for Jupiters is likely tied to a
large astrophysical false positive rate~\citep{santerne_sophie_2012}.
Going from Neptunes to Earths, the data suggest a weak increase in the
detected planet companion fraction, perhaps of a few percent.  It will
be interesting to see whether this effect is borne out by further
observations.}

\paragraph{Connection between high-resolution imaging and occurrence rates}

Occurrence rate calculations are beginning to incorporate the results
of high-resolution imaging surveys.  For example, recent studies have
reduced contamination in the numerator of the occurrence rate by
using~\citet{furlan_kepler_2017}'s catalog to test the effects of
removing KOI hosts with known companions
(\citealt{fulton_california-_2017}; \citealt{petigura_CKS_2017}).
This is a good first step, but the denominator remains uncorrected.

Even if the multiplicity of every star that {\it Kepler} surveyed were
known, the problem would not be solved. In this scenario, it would be
possible to tabulate planet occurrence for single and binary star
systems separately.  The number of planets per single star would
be known.  However the number of planets per primary and secondary
would still be convolved.  An observational approach to separating the
two populations might be to analyze the centroids, or the transit
durations (and thus host star densities) for a representative sample
of planets in binary systems.

A theory-driven solution might entail probabilistic population
inference, with a parametrized model for planet populations around
primary and secondary stars.  Such an approach might be expected to
find that the primary and secondary populations are different, since
single M dwarfs host more small planets than FGK
dwarfs~\citep{dressing_occurrence_2015}.  Given the difficulty of
writing a likelihood function in a survey with systematic
biases, one possible approach might be to forward-model the survey in
the framework of approximate Bayesian
computation~\citep[\textit{e.g.},][]{morehead_understanding_2016}.

\end{comment}

\begin{comment}
%LB: the below paragraph seems like it is just making caveats for no
%good reason. We've been explicit throughout that we have a
%restrictive set of assumptions...

  \paragraph{Practical effects: smooth detection efficiencies and finite SNR floors}
  To apply this study to real transit surveys, a few practical concerns
  become pressing.  One issue is that space-based missions with
  telemetric requirements often must select target stars.  Even if the
  selection process aims to create an SNR-limited survey,
  as~\citet{batalha_selection_2010} did for {\it Kepler}, uncertainties
  in stellar parameters and the modeled instrument performance can lead
  to deviations away from an SNR-limited stellar sample.

  Another concern is that in real transit surveys the detection
  efficiency is not a step function in SNR.  Transit survey pipeline
  detection efficiencies are usually smooth
  functions~\citep{christiansen_measuring_2016,fulton_california-_2017}.
  If we were to assume a smooth detection probability in SNR, we would
  need to appropriately weight our occurrence calculation for planets
  with low probabilities of being detected.

  The simplest way to avoid the conceptual complications of accounting
  for finite detection efficiency is to raise the SNR floor, to {\it
  e.g.,} ${\rm SNR} \approx 12$ for {\it Kepler}
  (\citealt{fulton_california-_2017}, Figure~5).  To a good
  approximation, this allows use of the boolean distinctions
  ``searchable'' and ``not searchable'' for signals of an observed depth
  (at fixed orbital period).  By making the survey strictly SNR-limited,
  rather than ``fuzzily SNR-limited'', the effects of unrecognized
  binaries become much easier to model.  Since the minimum SNR for
  detection is set arbitrarily\footnote{For example,
  {\it Kepler}'s threshold of ${\rm SNR}=7.1$ was set for there to be no
  more than one statistical false alarm across the full {\it Kepler}
  campaign~\citep{jenkins_tests_2002}. Equally valid would have been to
  insist on no more than $10^{-3}$ false alarms per survey, raising the
  detection floor.  }, this simplification would put occurrence rate
  studies on clearer conceptual ground, and facilitate the process of
  converting the apparent radius distributions into true radius
  distributions.

  Finally, our calculations ignored the fact that a transit survey's
  finite SNR floor could censor the apparent rate density. This would
  make planets with smaller apparent radii harder to detect.  Other than
  moving the small radii regime out of reach for that particular survey,
  this would not change any of our findings.

\end{comment}

\begin{comment}
%LB: the following are paragraphs from earlier discussion iterations

  To count the number of stars that
  were searched, further effort to understand the multiplicity
  statistics of KOI vs.~non-KOI stars is necessary. In particular, a
  non-KOI star high-resolution imaging effort could help estimate the
  systematic errors incurred by assuming all surveyed stars are single.
  An alternate approach would be to combine {\it Gaia} parallaxes with
  measured photometric fluxes and identify binary systems as those that
  are overluminous~\citep[\textit{e.g.},][]{widmark_inferring_2018}.

  \citet{ciardi_understanding_2015} studied the effects of stellar
  multiplicity on the planet radii derived from transit surveys.  They
  modeled the problem for {\it Kepler}\ objects of interest by
  matching a population of binary and tertiary companions to KOI
  stars, under the assumption that the KIC-listed stars were the
  primaries.  They then computed planet radius correction factors
  assuming that {\it Kepler}-detected planets orbited the primary or
  companion stars with equal probability (their Section~5).  Under
  these assumptions, they found that any given planet's radius is on
  average underestimated by a multiplicative factor of $\approx\!
  1.5$.  \citet{ziegler_robo-ao_2017} recently reported a similar mean
  radius correction factor for the multiple systems observed by the
  Robo-AO KOI survey, using the same assumption that detected planets
  are as likely to orbit the primary and the secondary.

  Of course, our model's assumptions (see the list in
  Section~\ref{sec:conclusion}) might not apply to the {\it Kepler}\
  dataset.  However if they are applicable, then for the $56\%$ of
  ``\texttt{CONFIRMED}'' KOIs\footnote{Exoplanet Archive;
  \cite{akeson_nasa_2013}; \url{exoplanetarchive.ipac.caltech.edu}} with
  apparent radii $>2r_\oplus$, whenever high-resolution imaging
  discovers a binary companion in a system that hosts a detected
  transiting planet, the planet is much more likely to orbit the
  primary.

  \paragraph{Independent approaches for estimating binarity's effects}
  T. Barclay et al.\! (in preparation) have performed the exercise of
  taking stars selected by the {\it Kepler}\ team, pairing them with a
  population of secondaries, injecting a realistic distribution of
  planet radii, and then comparing the inferred occurrence rates with
  the true ones.  In their model, they find that binarity leads to an
  inferred rate of Earth-sized planets $\approx 10\%$ less than the
  true rate.  In our Model \#3, if all $Z_i$'s are equal (a plausible
  assumption in the lack of evidence to the contrary), the
  underestimate is by a comparable 16\%.
\end{comment}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{comment}
\section{Conclusion}
% JNW: Material seems redundant.
  This study presented a framework for estimating the impact of
unresolved binaries on transit survey occurrence rates. From
the order-of-magnitude argument presented in
Section~\ref{sec:simplest}, we showed that for Sun-like stars in the
local neighborhood, typical twin binary fractions yield apparent
occurrence rates that are well within a factor of two of the true
rates for single stars.

We then derived a general formula for the apparent rate density
inferred by observers who ignore binarity
(Equation~\ref{eq:general_Gamma_a}).  As input, this equation requires
a volume-limited mass ratio distribution, and the true rate densities
for planets around singles, primaries, and secondaries.  The
assumptions that enable this approach are as follows.
\begin{enumerate}
    \item The transit survey is SNR-limited
        (Equation~\ref{eq:S_N_thresh}).
    \item The observers know the true properties of the single
        and primary stars, and assign all unresolved binaries the
        properties of the primary.
    \item There are two functions, $L(M)$ and $R(M)$, that specify a
        star's luminosity and radius in terms of its mass.
\end{enumerate}

The interpretation of Equation~\ref{eq:general_Gamma_a} is that the
apparent rate density is a weighted sum of the rate densities for
single, primary, and secondary stars. The weights depend on the
relative numbers of binary and single stars in the searchable volumes,
which are affected by both dilution of transit signals and a Malmquist
bias.  Both effects should be considered in Monte Carlo simulations of
transit surveys~\citep[\textit{e.g.},][]{
  bakos_hatsouth:_2013,sullivan_transiting_2015, gunther_new_2017}.

Applying Equation~\ref{eq:general_Gamma_a}, we then showed that:
\begin{itemize}
%
\item If single, primary, and secondary stars all host the same number
  of planets per star, and the planet radius distribution is a power
    law with the exponent reported by~\citet{howard_planet_2012}, then
    binarity influences apparent planet occurrence rates at the few
    percent level. This applies for apparent radii from $2R_\oplus$ to
    $17R_\oplus$ (Section~\ref{sec:model_2}).
%
\item Assuming a broken-power law planet radius distribution, with the
  \citet{howard_planet_2012} exponent above $2R_\oplus$ and a constant
    occurrence below $2R_\oplus$, the rate density for radii below
    $2R_\oplus$ is overestimated. The magnitude of the error, $\Delta
    \Gamma_0 = |\Gamma_0 - \Gamma\a|/\Gamma_0$, can easily reach
    $50\%$ (Figure~\ref{fig:occ_rate_model_3_log}).  Although this is
    smaller than current systematic uncertainties on the occurrence
    rates of Earth-sized planets, this implies that binarity could
    eventually become an important component of the $\eta_\oplus$
    error budget.
%
\item Binarity blurs sharp features in the radius distribution
  (Figure~\ref{fig:model_4}), to a degree that could affect precise
    measurements of the depth, width, and slope of
    the~\citet{fulton_california-_2017} radius gap.
    %in the event that
    %planets are not carefully vetted with high-resolution imaging.
%
\item Binarity does not lead to smaller apparent HJ occurrence rates
  (Figure~\ref{fig:gaussian_HJ}).  This assumes a Gaussian planet
    radius distribution, similar to that reported
    by~\citet{petigura_CKS_2017}.
%
\item Detected planets with apparent radii greater than $0.5R_\oplus$
  that are revealed by high-resolution imaging surveys to exist in
    binaries are more likely to orbit the primary
    (Figure~\ref{fig:frac_model_3}).
%
\item Near the break in occurrence as a function of planet
  radius at $2R_\oplus$, the fraction of detected planets with
    binary companions should increase by roughly $5$~to~$10\%$
    (Figure~\ref{fig:frac_model_3}).
\end{itemize}

These results are only strictly applicable for idealized transit
surveys meeting the criteria mentioned above.  For real transit
surveys, although our approach is only suggestive, we hope that it
provides a useful estimate for the systematic errors that can be
incurred by ignoring binarity in calculations of planetary occurrence
rates.
\end{comment}

\acknowledgements{
It was a pleasure discussing this study with T.~Barclay, W.~Bhatti,
J.~Christiansen, F.~Dai, and T.~Morton.  This work made use of NASA's
Astrophysics Data System Bibliographic Services.
%
\newline
%
\software{\texttt{numpy}~\citep{walt_numpy_2011}, 
\texttt{scipy}~\citep{jones_scipy_2001}, 
\texttt{matplotlib}~\citep{hunter_matplotlib_2007}, 
\texttt{pandas}~\citep{mckinney-proc-scipy-2010},
\texttt{IPython}~\citep{perez_2007}
}
}

\newpage
\appendix
\section{Alternative derivation of apparent rate density}
\label{sec:appendix}

Here we present an alternative derivation of
Equation~\ref{eq:general_Gamma_a}.  We also generalize a bit further,
by allowing the occurrence rate density to depend on stellar mass $M$
as well as planet radius $r$.  We begin by writing the apparent rate
density $\Gamma\a$ as (in analogy to
Equation~\ref{eq:occ_rate_simple})
\begin{equation}
    \Gamma\a(r\a,M\a) = \frac{n_{\rm det}(r\a,M\a)}{N_\star(r\a,M\a)}
    \, \frac{1}{p_{\rm tra}(M\a)},
\label{eq:apparent_rate_density}
\end{equation}
where $n_{\rm det}(r\a,M\a)$ is the number of detected planets per
star with radius $r\a$ orbiting stars of apparent mass $M\a$, per unit
$r\a$ and $M\a$.  This collection of planets with the same apparent
properties actually includes planets with a variety of true sizes and
host star masses.  We may write $n_{\rm det}(r\a, M\a)$ as the
convolution of the true occurrence rate density, $\Gamma(r, M)$; and
$\mathcal{N}_\star(r\a, M\a; r, M)$, the number of searchable stars
per unit $(r\a, M\a)$ for which a planet with true properties $(r, M)$
would appear to be a planet with apparent properties $(r\a, M\a)$:
\begin{align}
    n_{\rm det}(r\a, M\a) &=
    \sum_{i=0}^2 n_{\rm det}^i(r\a, M\a) \\
    &=
    \sum_{i=0}^2 \int \mathrm{d}r \!\int\! \mathrm{d}M \,
    \mathcal{N}_\star^i(r\a,M\a; r,M)
    \,\Gamma_i(r,M) \, p_{\rm tra}(M),
    \label{eq:n_det}
\end{align}
where $i$ specifies the true type of host star, with 0 for single
stars, 1 for primary stars, and 2 for secondary stars.  We now proceed
to calculate $\mathcal{N}_\star^i(r\a,M\a; r,M)$.

\paragraph{Single stars}
For single stars,
\begin{equation}
    \mathcal{N}_\star^0(r\a,M\a; r,M)
    =\hat{\delta}(r\a-r)\hat{\delta}(M\a-M) N_\star^0(r,M),
\end{equation}
where $\hat{\delta}$ is the Dirac delta function.  Therefore,
\begin{equation}
    n_{\rm det}^0(r\a, M\a)=N_\star^0(r\a,M\a)\, \Gamma_0(r\a, 
    M\a) \, p_{\rm tra}(r\a, M\a).
\end{equation}

\paragraph{Primary stars}
For primary stars in binary systems, 
\begin{equation}
    \mathcal{N}_\star^1(r\a, M\a; r, M)
    = \int_0^1 
      \mathrm{d}q\,f(q)\,\mathcal{N}_{{\rm s}, q}^1(r\a, M\a, q; r, M),
    \label{eq:fancyN_s_1}
\end{equation}
where $f(q)$ is the binary mass ratio distribution in a volume-limited sample
of stars.  Since we are assuming that the observer will perceive a binary to be a single
star with the same properties as the primary star,
\begin{equation}
    \mathcal{N}_{{\rm s},q}^1(r\a, M\a, q; r, M)
    \propto
    \hat{\delta}(M\a-M).
\end{equation}
In this case, $\mathcal{N}_{{\rm s},q}^1$ is non-zero only at
$r\a=R\a\sqrt{\delta}$, and the observed transit depth is
\begin{equation}
    \delta
    = \left[{r\over R(M\a)}\right]^2\, {L(M\a) \over L_{\rm tot}(M\a, q)}
    \equiv \left[{r\over R(M\a)}\right]^2\, 
    \frac{1}{\mathcal{D}_1^2}
    \label{eq:delta_obs_primaries} 
\end{equation}
where
\begin{equation}
    L_{\rm tot}(M\a,q) = L(M\a) + L(qM\a),~~\mathcal{D}_1=\sqrt{L_{\rm tot}(M\a, q) \over L(M\a)}.
\end{equation}
The normalization of $\mathcal{N}_{{\rm s},q}^1$ is given by the
number of binaries that are searchable for a signal $\delta$
and that have mass ratio $q$:
\begin{equation}
    N_\star^0(r\a, M\a)\,
    {n\b\over n\s}\left[L_{\rm tot}(M\a, q) \over L(M\a)\right]^{3/2}.
    \label{eq:primary_normalization}
\end{equation}
%where the argument for the number of searchable single stars,
%$N_\star^0$, could also be expressed as $(\delta, L(M\a))$.
Applying Equation~\ref{eq:primary_normalization},
\begin{align}
    \mathcal{N}_{{\rm s}, q}^1(r\a, M\a, q; r, M)
    &=N_\star^0(r\a, M\a)
    \, {n\b\over n\s} \, \mathcal{D}_1^3
    \,\hat{\delta} \left[r\a-\frac{r}{\mathcal{D}_1}\right]\hat{\delta}(M\a-M).
    \label{eq:N_s_1}
\end{align}

\paragraph{Secondary stars}
In this case, $M=qM_1=qM\a$, so
\begin{equation}
    \mathcal{N}_{{\rm s},q}^2(r\a, M\a, q; r, M)
    \propto \hat{\delta}\left(M\a-{M\over q}\right).
    \label{eq:fancyN_s_2_prop}
\end{equation}
Again $\mathcal{N}_\star^2$ is non-zero only at
$r\a=R\a\sqrt{\delta}$, but this time
\begin{equation}
      \delta =
      \left[{r\over R(qM\a)}\right]^2 \,
      {L(qM\a) \over L_{\rm tot}(M\a, q)}
      \equiv
      \left[{r\over R(M\a)}\right]^2\,
      \frac{1}{\mathcal{D}_2^2},
\end{equation}
where
\begin{equation}
  \mathcal{D}_2
  =
  \frac{R(qM\a)}{R(M\a)}
  \sqrt{ \frac{L_{\rm tot}(M\a,q)}{L(qM\a)} }.
\end{equation}
The normalization remains the same as the previous case~--~we are
counting the searchable stars at a given observed depth $\delta$, and
the total luminosity of the binary is the same.  Thus,
\begin{equation}
    \mathcal{N}_\star^2(r\a, M\a; r, M)
    =\int_0^1 \mathrm{d}q~f(q)\,\mathcal{N}_{{\rm s}, q}^2(r\a, M\a, q; r, M),
\end{equation}
where
\begin{align}
    \mathcal{N}_{{\rm s},q}^2(r\a, M\a, q; r, M)
    =N_\star^0(r\a, M\a) \, \frac{n\b}{n\s} 
    \, \mathcal{D}_1^3
    \, \hat{\delta} \left(r\a-\frac{r}{\mathcal{D}_2}\right)
    \hat{\delta} \left(M\a-{M\over q}\right).
    \label{eq:mathcal_N_s_2}
\end{align}

\paragraph{Number of detected planets}
Inserting the results into Equation~\ref{eq:n_det}, we find
\begin{align}
    \notag
    n_{\rm det}^0(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^0(r\a, M\a; r, M)
    \, \Gamma_0(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, \Gamma_0(r\a, M\a) \, p_{\rm 
        tra}(M\a),\\
    \notag
    n_{\rm det}^1(r\a, M\a)
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^1(r\a, M\a; r, M)
    \, \Gamma_1(r, M) \, p_{\rm tra}(M)\\
    &=N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \,
    \frac{n\b}{n\s}
    \int \mathrm{d}q \,
         \mathcal{D}_1^3 f(q)\cdot
         \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a),\\
    n_{\rm det}^2(r\a, M\a)
    \notag
    &=\int\mathrm{d}r\,\mathrm{d}M~\mathcal{N}_\star^2(r\a, M\a; r, M)
    \, \Gamma_2(r, M) \, p_{\rm tra}(M)\\
    &=
    N_\star^0(r\a, M\a)\, p_{\rm tra}(M\a) \, \frac{n\b}{n\s}
    \int {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\cdot
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}.
    \label{eq:n_det_2}
\end{align}

Finally, combining these results, the
apparent occurrence rate density
\begin{align}
    \Gamma\a(r\a,M\a) &= 
    \frac{1}{N_\star(r\a,M\a) p_{\rm tra}(r\a,M\a)} \,
    \sum_{i=0}^2 n_{\rm det}^i (r\a,M\a)
\end{align}
becomes
\begin{align}
    \notag
    \Gamma\a(r\a,M\a) &= \frac{1}{1+\mu} \,
    \left\{ \Gamma_0(r\a, M\a)+ 
    \frac{n\b}{n\s}
    \left[ \int_0^1 \mathrm{d}q \,
           \mathcal{D}_1^3 f(q)\cdot
           \mathcal{D}_1 \Gamma_1(\mathcal{D}_1r\a, M\a)\,
    \right.   
    \right. \\
    & \quad\quad\quad\quad\quad \left.\left.
    +\int_0^1 {\rm d}q\, 
         \mathcal{D}_1^3 f(q)\cdot q \mathcal{D}_2
         \Gamma_2(\mathcal{D}_2r\a, qM\a)\cdot
         \frac{R(qM\a)}{R(M\a)} q^{-1/3}
    \right] \right\}.
\end{align}
%We validate this equation in limits where it is possible to write down
%the answer (e.g., Equation~\ref{eq:formula_twin_binary}), and also
%against a Monte Carlo realization of the twin binary models
%(Sections~\ref{sec:model_1} and~\ref{sec:model_2}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\input{hj_rates.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newpage
\bibliographystyle{yahapj}                            
\bibliography{bibliography}

\end{document}