### 2018-04-13 paper/review/response_1.txt
add response to referee comments

Dear Dr. Sigurdsson,

We thank you for organizing the review of our paper, and the referee
for constructive criticism.  We have considered the referee's remarks
carefully and revised our manuscript accordingly.

Below, we have reproduced the referee's report, along with our
point-by-point responses. A list of changes is also appended to the
manuscript.

Sincerely,

Luke Bouma, Kento Masuda, and Joshua Winn.

********************************************************************************

Referee comment:

Section 2, end of paragraph 2, For each of the 4 points regarding missed
transit signals, how often do they occur? (could add references).

  Response:

  We cannot identify "4 points regarding missed transit signals" in section 2,
  end of paragraph 2.

  Assuming that the referee is referring to points #1, #2, #3, enumerated at
  the end of section 2, paragraph 4:

  We do not know how often each of these effects happen -- the goal of this
  section is simply to introduce the existence of these sources of uncertainty.
  Could the referee please clarify their request, or perhaps give some indication
  of which references we might add?

********************************************************************************

Referee comment:

Section 3.1, first sentence, "Since the effects of binarity are most pronounced
when the two stellar components are similar" <- please also point to the place
in the text where this is explained.

  Response:

  The original idea was that binarity produces the largest errors in inferred
  planetary radii when the two stellar components are similar.  However, this
  is only true if the detected planet orbits the primary.  If it orbits the
  secondary, it is false.

  We have accordingly reworded this sentence.

********************************************************************************

Referee comment:

Section 3.2, following equations 4, you may want to change "stars" to "host
star".

  Response:

  Equation 4 is in Section 3.1, and the word "stars" does not appear around
  that equation in an immediately obvious way.

  Perhaps the referee meant around equation 14 of section 3.2?

  We have not taken any action on this point.

********************************************************************************

Referee comment:

Section 4.1, when exploring the Howard et al. planet radius distribution, how
sensitive are the results to the choice of gamma? Do mathematical artifacts
here make the results insensitive to the choice of the value of gamma?

  Response:

  Consider the cases of γ=-2.5, and γ=-3.5. (Well outside the power laws needed
  to fit Howard et al's results).  Applying Equations 28 and 30, we find:

                   γ=-2.5      γ=-2.92      γ=-3.5
  --------------------------------------------------
  Equation 28
  Γ_a/Γ_0           1.023        1.003      0.980

  Equation 30 
  Γ_a/Γ_0           1.095        1.048      0.993

  We have added a sentence summarizing this result.

********************************************************************************

Referee comment:

Section 4.4, end of paragraph 1, Add Lopez and Fortney citation.

  Response:

  We have corrected this omission.  In addition to referencing the work by
  Lopez & Fortney 2013, we now also refer to that by Owen & Wu 2013.

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, "radius distribution is assumed to have a complete
absence of planets with sizes between 1.5 and 2R_Earth", might clarify that the
"true" radius distribution is assumed to have this feature after fitting the
apparent distribution.

  Response:

  We have clarified the wording in this paragraph, in a manner that should
  clarify the distinction between Fulton+17's apparent distribution, and the true
  distribution that we assume for this numerical experiment.

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, More discussion of the spurious feature below 1.5
R_Earth would be helpful.

  Response:

  The peak in apparent occurrence rate density between 1.3 and 1.5 R_earth
  (solid blue line relative to dotted light blue line) is the contribution from
  secondaries with true radii greater than 2R_earth.

  We have added a comment to this effect in the caption of Figure 6. We have
  also increased the resolution of the grid in planet radius, to emphasize that
  this is a feature, rather than a bug.

  [
  OPTIONAL:
  As an aside, consider the numerical experiment of a single planet size, and a
  distribution of binary light ratios. I.e., imagine the true occurrence rate
  density is a delta function around r_p. The resulting apparent occurrence rate
  density would have contributions from primaries between r_a = r_p/sqrt(2) and
  r_a = r_p.  It would have contributions from secondaries from (roughly) r_a =
  0 up to r_p/sqrt(2). We ran this numerical experiment, and though did not
  mention it in the manuscript, we found that the contribution from secondaries
  peaked at r_a = r_p/sqrt(2).

  This delta function peak is analogous to what is seen in Figure 6.
  ]

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, "may be even more devoid of planets than it appears",
might want to emphasize again here that this conclusion is very dependent on
the true distribution model plugged into the formalism.

  Response:

  We agree that the apparent rate density distribution is very dependent on
  the assumed true rate density distribution. However, we have opted not to
  change our original wording, which we believe accurately states our
  conclusion.

  Our conclusion here is that for an assumed true rate density distribution
  that roughly matches observations, there is no way for binarity to
  cause underestimates of the rate density inside a "dip" of planet occurrence.

  Restated, if there is a "gap" in the true rate density distribution,
  the apparent rate density distribution will always fill it in.

  Thus, if a gap is observed in an apparent distribution, the true distribution
  must have a gap at least as deep, if not deeper.

********************************************************************************

Referee comment:

Section 5, Hot Jupiter paragraph, states that result of this study find that
unresolved binaries likely result in an overestimated of hot-Jupiter occurrence
rates, implied by the Fulton gap model in Section 4.4. Previously, in section
4.2, paragraph 3, it states the occurrence rate density of planets greater than
8 Earth radii are underestimated. The abstract just says net systematic error
of order 5%. It would help to include a table describing the different tests in
this study, and how parameters are varied (how knobs are turned).

  Response:

  In Section 5, Hot Jupiter paragraph, we erroneously referred to Section 4.4;
  this should have been a reference to Section 4.5 (the gaussian radius
  distribution model).  We have corrected the error.

  We have now also added a table summarizing the case studies, as well as
  the parameters that are varied.

********************************************************************************

Referee comment:

- Recommend using the term "uncertainty" instead of "errors" .

  Response:

  We have changed our terminology accordingly in many places throughout the
  manuscript (see list of changes).

********************************************************************************

Referee comment:

- May want to comment on parameterizing the occurrence rate in terms of planet
  radius and insolation flux. The semi major axis will depend on which star in
  the binary the planet orbits.

- May want to comment on paper by Danley Hsu: Occurrence rates from Kepler
  using approximate Bayesian computation. You could speculate on what their
  data generating process would yield when plugged into this formalism.

- May want to comment on the parameter space and telescopes for resolving
  binaries of various separations, etc. and the probability of being able to
  resolve binaries with followup.

  Response:

  We have added a "Caveats and Outlook" paragraph that touches on all three of
  these issues.