### 2018-04-13 paper/case_study_toc.tex

%% \begin{deluxetable}{} command tell LaTeX how many columns
%% there are and how to align them.
\begin{deluxetable}{cccc}

%% Keep a portrait orientation

%% Over-ride the default font size
%% Use Default (12pt)

%% Use \tablewidth{?pt} to over-ride the default table width.
%% If you are unhappy with the default look at the end of the
%% *.log file to see what the default was set at before adjusting
%% this value.

%% This is the title of the table.
\caption{Case study table of contents.}
\label{tab:case_study_toc}

%% This command over-rides LaTeX's natural table count
%% and replaces it with this number.  LaTeX will increment 
%% all other tables after this table based on this number
\tablenum{1}

%% The \tablehead gives provides the column headers.  It
%% is currently set up so that the column labels are on the
%% top line and the units surrounded by ()s are in the 
%% bottom line.  You may add more header information by writing
%% another line between these lines. For each column that requries
%% extra information be sure to include a \colhead{text} command
%% and remember to end any extra lines with \\ and include the 
%% correct number of &s.
\tablehead{
  \colhead{Section} &
  \colhead{True rate density} &
  \colhead{Apparent rate density} & 
  \colhead{Applicable where?}
}

%% All data must appear between the \startdata and \enddata commands
\startdata
  \ref{sec:model_2} & power-law & Eqs.~\ref{eq:power_law_correction}
  \&~\ref{eq:powerlaw_vary_binary} & 2-17\,$R_\oplus$; no gap; no HJs
  \\
  \ref{sec:model_3} \& \ref{subsec:whichstar} & broken power-law &
  Fig.~\ref{fig:occ_rate_model_3_log} & 0-17\,$R_\oplus$; 
  no gap; no HJs
  \\
  \ref{sec:further_models} & broken power-law + gap & Fig.~\ref{fig:model_4} & 
  0-17\,$R_\oplus$; has gap; no HJs
  \\
  \ref{sec:hjs} & gaussian & Fig.~\ref{fig:gaussian_HJ} & only hot Jupiters
  \\
\enddata

%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
%% or \tablecomments{text} between the \enddata and 
%% \end{deluxetable} commands

%% General table comment marker
\tablecomments{
    Aside from varying the true rate density, within each case study we also 
    vary the number of planets orbiting secondary stars, relative to the 
    number orbiting single stars. We use the following notation throughout.
    $\alpha$---power-law exponent in stellar mass-luminosity relation, 
    $L\propto M^\alpha$;
    $\beta$---power-law exponent in volume-limited binary mass-ratio 
    distribution, $f(q)\propto q^\beta$;
    $\gamma$---power-law exponent in true rate density above 2\,$R_\oplus$, 
    $\Gamma(r)\propto r^\gamma$.
}

\end{deluxetable}