### 2018-04-18 paper/review/response_1.txt
update response per JNW comments

Dear Dr. Sigurdsson,

We thank you for organizing the review of our paper, and the referee
for constructive criticism.  We have considered the referee's remarks
carefully and revised our manuscript accordingly.

Below, we have reproduced the referee's report, along with our
point-by-point responses. A list of changes is also appended to the
manuscript.

Sincerely,

Luke Bouma, Kento Masuda, and Joshua Winn.

********************************************************************************

Referee comment:

Section 2, end of paragraph 2, For each of the 4 points regarding missed
transit signals, how often do they occur? (could add references).

  Response:

  We assume the referee is referring to paragraph 3.  All of these
  effects are quite common, both in the Kepler survey and ground-based
  surveys, though we are not aware of any source that would allow us
  to estimate the probabilities.  Since these "pitfalls" are not the
  topic of this paper, we do not want to attempt such a calculation
  here.

********************************************************************************

Referee comment:

Section 3.1, first sentence, "Since the effects of binarity are most pronounced
when the two stellar components are similar" <- please also point to the place
in the text where this is explained.

  Response:

  We decided to omit this sentence.  The original idea was that
  binarity produces the largest errors in inferred planetary radii
  when the two stellar components are similar.  However, this is only
  true if the detected planet orbits the primary.  If it orbits the
  secondary, it is false.

********************************************************************************

Referee comment:

Section 3.2, following equations 4, you may want to change "stars" to "host
star".

  Response:

  Equation 4 is in Section 3.1 (not Section 3.2) and is not followed
  by a sentence with the word "stars". Perhaps the referee is
  referring to equation 14?  We do not understand why it would be an
  improvement here to replace "stars" with "host star", so we have not
  taken any action.

********************************************************************************

Referee comment:

Section 4.1, when exploring the Howard et al. planet radius distribution, how
sensitive are the results to the choice of gamma? Do mathematical artifacts
here make the results insensitive to the choice of the value of gamma?

  Response:

  We addressed this point by considering the cases of γ=-2.5, and
  γ=-3.5, which bracket the result from Howard et al. and are well
  outside the quoted uncertainty range.  Applying Equations 28 and 30,
  we find:

                   γ=-2.5      γ=-2.92      γ=-3.5
  --------------------------------------------------
  Equation 28
  Γ_a/Γ_0           1.023        1.003      0.980

  Equation 30 
  Γ_a/Γ_0           1.095        1.048      0.993

  We have added a sentence summarizing this result.

********************************************************************************

Referee comment:

Section 4.4, end of paragraph 1, Add Lopez and Fortney citation.

  Response:

  We have corrected this omission, and also added a citation to Owen &
  Wu (2013).

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, "radius distribution is assumed to have a
complete absence of planets with sizes between 1.5 and 2R_Earth",
might clarify that the "true" radius distribution is assumed to have
this feature after fitting the apparent distribution.

  Response:

  We have tried to clarify this passage, to distinguish between the
  apparent rate distribution measured by Fulton et al. (2017) and the
  distribution assumed in our numerical experiment.

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, More discussion of the spurious feature below 1.5
R_Earth would be helpful.

  Response:

  The peak in apparent occurrence rate density between 1.3 and 1.5
  R_earth (solid blue line relative to dotted light blue line) is the
  contribution from secondaries with true radii greater than 2R_earth.
  This is now stated in the caption of Figure 6.  We have also
  increased the resolution of the grid in planet radius, to emphasize
  that this is a feature, rather than a bug.

********************************************************************************

Referee comment:

Section 4.4, paragraph 2, "may be even more devoid of planets than it appears",
might want to emphasize again here that this conclusion is very dependent on
the true distribution model plugged into the formalism.

  Response:

  We have added a few sentences to make our point more clearly:
  "The quantitative results of this numerical experiment obviously
  depend on the input assumptions.  The more general point is that
  whenever there is a gap-like feature in the true rate density
  distribution, the effects of binarity will tend to fill it in.  If a
  gap is observed in an apparent distribution, the true distribution
  must have a gap at least as deep, if not deeper."

********************************************************************************

Referee comment:

Section 5, Hot Jupiter paragraph, states that result of this study find that
unresolved binaries likely result in an overestimated of hot-Jupiter occurrence
rates, implied by the Fulton gap model in Section 4.4. Previously, in section
4.2, paragraph 3, it states the occurrence rate density of planets greater than
8 Earth radii are underestimated. The abstract just says net systematic error
of order 5%. It would help to include a table describing the different tests in
this study, and how parameters are varied (how knobs are turned).

  Response:

  We see now that in Section 5 (Hot Jupiter paragraph) we erroneously
  referred to Section 4.4; this should have been a reference to
  Section 4.5 (the gaussian radius distribution model).  We have
  corrected the error.

  We have also followed the referee's advice and added a table
  summarizing the case studies, as well as the parameters that are
  varied.

********************************************************************************

Referee comment:

- Recommend using the term "uncertainty" instead of "errors" .

  Response:

  We appreciate the suggestion but we prefer to leave the text as it
  is.  The term "systematic error" is in more common use than
  "systematic uncertainty", by about a factor of 5 according to Google
  Ngram viewer.

********************************************************************************

Referee comment:

- May want to comment on parameterizing the occurrence rate in terms of planet
  radius and insolation flux. The semi major axis will depend on which star in
  the binary the planet orbits.

  Response: We have added a "Caveats and Outlook" paragraph that
  touches on all three of these issues.

- May want to comment on paper by Danley Hsu: Occurrence rates from
  Kepler using approximate Bayesian computation. You could speculate
  on what their data generating process would yield when plugged into
  this formalism.

  Response: The work by Hsu et al. does look interesting, but since we
  lack experience with their technique we do not wish to speculate on
  how to modify it to accommodate binarity.

- May want to comment on the parameter space and telescopes for resolving
  binaries of various separations, etc. and the probability of being able to
  resolve binaries with followup.

  Response: Given that there are so many possible approaches and
  instruments for imaging, RV, and astrometry, a quantitative analysis
  of this issue would be beyond the scope of this article.  In the
  revision, we have simply acknowledged that the combination of Gaia
  and TESS data will lessen the severity of the problem, compared to
  the situation with Kepler.

********************************************************************************

Other:

* We have changed the ylabel in Figures 4b and 6b from "number of planets per
  star" to "number of planets per star per bin".

* We now mention Figure 2 in the text (cf. appended List of Changes)